---
title: "MECC Hydraulic Analysis"
author: "FlowWest"
date: "12 June 2023"
output: 
  github_document:
    default
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
ggplot2::theme_set(theme_classic())
source("R/xs.R")
```

## Import cross sections

```{r message=FALSE, warning=FALSE}
xs_2007 <- read_csv("data/xs_2007.csv") %>%
  prep_xs(sta = station_ft, elev = elevation_ft) 

xs_2007 %>% ggplot(aes(y = gse, x = sta)) + geom_point() + 
  ylab("Elevation (ft)") + xlab("Station (ft)") + ggtitle("2007 Baseline") + coord_fixed(ratio = 1)
```

```{r message=FALSE, warning=FALSE}
xs_2023 <- read_csv("data/xs_2023_simplified.csv") %>% #_simplified.csv") %>%
  prep_xs(sta = station_ft, elev = elevation_ft) 

xs_2023 %>% ggplot(aes(y = gse, x = sta)) + geom_point() + 
  ylab("Elevation (ft)") + xlab("Station (ft)") + ggtitle("2023 Proposed") + coord_fixed(ratio = 1)
```

```{r message=FALSE, warning=FALSE}
ggplot() + 
  geom_line(data = xs_2007, aes(x = sta, y = gse, color = "2007 Baseline")) +
  geom_line(data = xs_2023, aes(x = sta, y = gse, color = "2023 Proposed")) +
  ylab("Elevation (ft)") + xlab("Station (ft)") + theme(legend.position="top", legend.title=element_blank())  + coord_fixed(ratio = 1)
```

## Calculate depth-discharge rating curves 

```{r message=FALSE, warning=FALSE}

model_slope <- (38.6-38.0) / 28.65

model_roughness <- 0.035 # n

print(c("model_slope" = model_slope, "model_roughness" = model_roughness))
```

*Future improvement to script should allow different slope and roughness for each scenario.*

```{r message=FALSE }
depth_vs_discharge_2007 <- xs_2007 %>% 
  calculate_rating_curve(slope = model_slope,
                         mannings_n = model_roughness) # 0.030) #model_roughness)
depth_vs_discharge_2007 %>% 
  ggplot(aes(y = max_depth, x = discharge_cfs)) + 
  geom_line() + ylab("Depth (ft)") + xlab("Discharge (cfs)") + ggtitle("2007 Baseline")
```

```{r message=FALSE, warning=FALSE}
depth_vs_discharge_2023 <- xs_2023 %>% 
  calculate_rating_curve(slope = model_slope,
                         mannings_n = model_roughness)
depth_vs_discharge_2023 %>% 
  ggplot(aes(y = max_depth, x = discharge_cfs)) + 
  geom_line() + ylab("Depth (ft)") + xlab("Discharge (cfs)") + ggtitle("2023 Proposed") 
```

```{r message=FALSE, warning=FALSE}
ggplot() +
  geom_line(data = depth_vs_discharge_2007, aes(y = max_depth, x = discharge_cfs, color = "2007 Baseline")) +
  geom_line(data = depth_vs_discharge_2023, aes(y = max_depth, x = discharge_cfs, color = "2023 Proposed")) + 
  ylab("Depth (ft)") + xlab("Discharge (cfs)") + theme(legend.position="top", legend.title=element_blank())
```
```{r message=FALSE, warning=FALSE}
ggplot() +
  geom_line(data = depth_vs_discharge_2007, aes(y = water_surface_elevation, x = discharge_cfs, color = "2007 Baseline")) +
  geom_line(data = depth_vs_discharge_2023, aes(y = water_surface_elevation, x = discharge_cfs, color = "2023 Proposed")) + 
  ylab("Water Surface Elevation (ft)") + xlab("Discharge (cfs)") + theme(legend.position="top", legend.title=element_blank())
```

*Future improvement to script should include functions for the following analyses to avoid duplication.*

## Q100

### Interpolate water surface for selected discharge

```{r message=FALSE, warning=FALSE}
model_discharge <- 3980 # cfs
```

```{r message=FALSE, warning=FALSE}
wse_2007 <- depth_vs_discharge_2007 %>% interpolate_rating_curve(model_discharge)
print(c("wse 2007" = wse_2007))
```

```{r message=FALSE, warning=FALSE}
wse_2023 <- depth_vs_discharge_2023 %>% interpolate_rating_curve(model_discharge)
print(c("wse 2023" = wse_2023))
```

```{r message=FALSE, warning=FALSE}
xs_2007 %>% calc_xs(wse_2007)
```

```{r message=FALSE, warning=FALSE}
xs_2023 %>% calc_xs(wse_2023)
```

### Plot outputs

```{r message=FALSE, warning=FALSE}
xs_2007 %>% arrange(sta) %>%
      mutate(wse = case_when(wse_2007 > gse ~ wse_2007)) %>%
      ggplot(aes(x = sta)) + 
      geom_line(aes(y = gse)) + 
      geom_line(aes(y = wse)) + 
      xlab("Station (ft)") + ylab("Elevation (ft)") + ggtitle("2007 Baseline") + coord_fixed(ratio = 1)
```
```{r message=FALSE, warning=FALSE}
xs_2023 %>% arrange(sta) %>%
      mutate(wse = case_when(wse_2023 > gse ~ wse_2023)) %>%
      ggplot(aes(x = sta)) + 
      geom_line(aes(y = gse)) + 
      geom_line(aes(y = wse)) + 
      xlab("Station (ft)") + ylab("Elevation (ft)") + ggtitle("2023 Proposed") + coord_fixed(ratio = 1)
```

```{r message=FALSE, warning=FALSE}
ggplot() + 
  geom_line(data = xs_2007, aes(x = sta, y = gse, color = "2007 Baseline", linetype = "Terrain")) +
  geom_line(data = xs_2023, aes(x = sta, y = gse, color = "2023 Proposed", linetype = "Terrain")) +
  geom_line(data = xs_2007, aes(x = sta, y = case_when(wse_2007 > gse ~ wse_2007), 
                                color = "2007 Baseline", linetype = "Water Surface")) + 
  geom_line(data = xs_2023, aes(x = sta, y = case_when(wse_2023 > gse ~ wse_2023), 
                                color = "2023 Proposed", linetype = "Water Surface")) + 
      xlab("Station (ft)") + ylab("Elevation (ft)") + theme(legend.position="top", legend.title=element_blank()) + coord_fixed(ratio = 1)
```

### Calculate design velocity based on Q100 flow

```{r message=FALSE, warning=FALSE}

velocity_2007 <- model_discharge / calc_xs(xs_2007, wse_2007)$cross_sectional_area
print(velocity_2007)

```

```{r message=FALSE, warning=FALSE}

velocity_2023 <- model_discharge / calc_xs(xs_2023, wse_2023)$cross_sectional_area
print(velocity_2023)

```

## Q2

### Interpolate water surface for selected discharge

```{r message=FALSE, warning=FALSE}
model_discharge <- 513 # cfs
```

```{r message=FALSE, warning=FALSE}
wse_2007_Q2 <- depth_vs_discharge_2007 %>% interpolate_rating_curve(model_discharge)
print(c("wse 2007" = wse_2007_Q2))
```

```{r message=FALSE, warning=FALSE}
wse_2023_Q2 <- depth_vs_discharge_2023 %>% interpolate_rating_curve(model_discharge)
print(c("wse 2023" = wse_2023_Q2))
```

```{r message=FALSE, warning=FALSE}
xs_2007 %>% calc_xs(wse_2007_Q2)
```

```{r message=FALSE, warning=FALSE}
xs_2023 %>% calc_xs(wse_2023_Q2)
```

### Plot outputs

```{r message=FALSE, warning=FALSE}
xs_2007 %>% arrange(sta) %>%
      mutate(wse = case_when(wse_2007_Q2 > gse ~ wse_2007_Q2)) %>%
      ggplot(aes(x = sta)) + 
      geom_line(aes(y = gse)) + 
      geom_line(aes(y = wse)) + 
      xlab("Station (ft)") + ylab("Elevation (ft)") + ggtitle("2007 Baseline") + coord_fixed(ratio = 1)
```
```{r message=FALSE, warning=FALSE}
xs_2023 %>% arrange(sta) %>%
      mutate(wse = case_when(wse_2023_Q2 > gse ~ wse_2023_Q2)) %>%
      ggplot(aes(x = sta)) + 
      geom_line(aes(y = gse)) + 
      geom_line(aes(y = wse)) + 
      xlab("Station (ft)") + ylab("Elevation (ft)") + ggtitle("2023 Proposed") + coord_fixed(ratio = 1)
```

```{r message=FALSE, warning=FALSE}
ggplot() + 
  geom_line(data = xs_2007, aes(x = sta, y = gse, color = "2007 Baseline", linetype = "Terrain")) +
  geom_line(data = xs_2023, aes(x = sta, y = gse, color = "2023 Proposed", linetype = "Terrain")) +
  geom_line(data = xs_2007, aes(x = sta, y = case_when(wse_2007_Q2 > gse ~ wse_2007_Q2), 
                                color = "2007 Baseline", linetype = "Water Surface")) + 
  geom_line(data = xs_2023, aes(x = sta, y = case_when(wse_2023_Q2 > gse ~ wse_2023_Q2), 
                                color = "2023 Proposed", linetype = "Water Surface")) + 
      xlab("Station (ft)") + ylab("Elevation (ft)") + theme(legend.position="top", legend.title=element_blank()) + coord_fixed(ratio = 1)
```

### Calculate design velocity based on Q2 flow

```{r message=FALSE, warning=FALSE}

velocity_2007_Q2 <- model_discharge / calc_xs(xs_2007, wse_2007_Q2)$cross_sectional_area
print(velocity_2007_Q2)

```

```{r message=FALSE, warning=FALSE}

velocity_2023_Q2 <- model_discharge / calc_xs(xs_2023, wse_2023_Q2)$cross_sectional_area
print(velocity_2023_Q2)

```

## Output design velocity tables

Calculate design velocities based on all modeled flood events from Q2 to Q100

```{r message=FALSE, warning=FALSE}
streamstats_result <- tribble(~return_interval, ~discharge,
                              2 ,   513,
                              5,   1320,
                              10,  1970,
                              25,  2780,
                              50,  3360,
                              100, 3980 ) 

velocities <- streamstats_result %>%
  mutate(xs_area_2007 = map_dbl(discharge, function(discharge) {
            depth_vs_discharge_2007 %>% 
            interpolate_rating_curve(., discharge) %>%
            calc_xs(xs_2007, .) %>% 
            .$cross_sectional_area
            }),
         xs_area_2023 = map_dbl(discharge, function(discharge) {
            depth_vs_discharge_2023 %>% 
            interpolate_rating_curve(., discharge) %>% 
            calc_xs(xs_2023, .) %>% 
            .$cross_sectional_area
            }),
         velocity_2007 = discharge / xs_area_2007,
         velocity_2023 = discharge / xs_area_2023
         )
velocities %>% knitr::kable()
```


```{r message=FALSE, warning=FALSE}
all_parms_2023 <- streamstats_result %>%
  mutate(xs_parms_2023 = map(discharge, function(discharge) {
            depth_vs_discharge_2023 %>% 
            interpolate_rating_curve(., discharge) %>% 
            calc_xs(xs_2023, .)})) |> unnest_wider(xs_parms_2023)
all_parms_2023 %>% knitr::kable()
```
