---
title: "Simple Cross Section Hydraulics using Manning's Equation"
output: 
  github_document:
    default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The **`xsmatic`** package contains generic functions for conducting Manning's equation hydraulic analyses based on surveyed cross section data:

* Clean up station-elevation table and densify coordinate points

* Calculate hydraulic geometry (cross-sectional area, wetted perimeter) for a given water surface elevation.

* Using Manning's *n* and a provided slope and roughness, calculate a rating curve of water surface elevation versus cross-sectional area, wetted perimeter, velocity, and discharge.

* Use a rating curve to return a water surface elevation for a given discharge value.

To install the package: 

```{r eval=FALSE, include=TRUE}

# authenticate with a github account within the FlowWest organization
usethis::use_git_config(user.name = "yourGithubUsername", user.email = "email@flowwest.com")
usethis::create_github_token() # copy the personal access token and paste below

# install private package
devtools::install_github(repo = "flowwest/xsmatic", 
                         auth_token = "TOKEN-HERE")
```

Load the package:

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(xsmatic)
```

Run via the interactive prompt (also published at [flowwest.shinyapps.io/xsmatic](https://flowwest.shinyapps.io/xsmatic)):

```{r eval=FALSE, include=TRUE}
xs_run_app()
```

Import cross sections:

```{r xs, message=FALSE, warning=FALSE}
xs_2007 <- read_csv("data-raw/xs_2007.csv") %>%
  xs_prep(sta = station_ft, elev = elevation_ft) 
xs_2023 <- read_csv("data-raw/xs_2023_simplified.csv") %>%
  xs_prep(sta = station_ft, elev = elevation_ft) 

xs_plot2(xs_2007, xs_2023)
```

Calculate rating curves:

```{r rc, message=FALSE, warning=FALSE}
rc_2007 <- xs_2007 %>% 
  xs_rating_curve(slope = 0.021, mannings_n = 0.035)
rc_2023 <- xs_2023 %>% 
  xs_rating_curve(slope = 0.021, mannings_n = 0.035)

xs_plot_rc2(rc_2007, rc_2023)
```

Calculate water surface elevation at given discharge

```{r wse, message=FALSE, warning=FALSE}
q100 <- 3980 # cfs
wse_2007 <- rc_2007 %>% xs_rc_interpolate(q100)
wse_2023 <- rc_2023 %>% xs_rc_interpolate(q100)

xs_plot2(xs_2007, xs_2023, wse_2007, wse_2023)
```

Calculate hydraulic parameters at multiple discharges

```{r eval-all, echo=TRUE, eval=FALSE, paged.print=TRUE}
xs_eval_all(xs = xs_2023, rc = rc_2023, 
            discharges = c("Q2"=513, "Q5"=1320, "Q10"=1970, "Q25"=2780, "Q50"=3360, "Q100"=3980))
```

```{r eval-all-kable, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
xs_eval_all(xs = xs_2023, rc = rc_2023, 
            discharges = c("Q2"=513, "Q5"=1320, "Q10"=1970, "Q25"=2780, "Q50"=3360, "Q100"=3980)) %>% knitr::kable()
```



